---
title: "COVID-19 Deaths By Vaccination Status"
author: "Jorge Valderrama"
date: "2023-12-23"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Importing Data
```{r}
#read data
setwd("~/Downloads")
USACOVID <- read.csv ("COVID_Data_R2.csv")

```
## Loading Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(gganimate)

```
##Demonstrating Untidy Data
```{r}
USACOVID
```
##Making Data Tidy
```{r}
tidyUSACOVID <-USACOVID %>% 
  pivot_longer(cols = -Month)

```
##Demonstrating Tidy Data
```{r}
tidyUSACOVID
```
##Loading lubridate package to change the “Month” column from character type variable to a date type variable
```{r}
library(lubridate)
```
##Changing variable type for “Month” variable
```{r}
tidyUSACOVID$Month <- mdy(tidyUSACOVID$Month)
```
##Demonstrating change in variable type
```{r}
tidyUSACOVID
```
##Creating a line graph to demonstrate differences in COVID-19 death rates among fully vaccinated and unvaccinated individuals over time.
```{r}
ggplot(data=tidyUSACOVID, aes(x=Month, y=value, group=name, color=name))+
  geom_line(size = 1.25)+
  ggtitle("Rates of COVID-19 Deaths by Vaccination Status in Ages 18 and Older")+
  theme_clean()+
  theme(axis.text.x=element_text(angle=60, hjust=1, face = "bold"))+
  theme(axis.title.y=element_text(size=14,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5, lineheight = 0.9))+
  theme(plot.title = element_text(face = "bold", size = 14))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(face = "bold", size = 10))+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits=c(0,5))+
  labs(y = "Deaths per 100,000 population", subtitle = "August 1, 2022 - March 26, 2023 (23 U.S. Jurisdictions)")+
  theme(plot.subtitle = element_text(hjust=0.5))+
  scale_color_hue(labels=c('Unvaccinated', 'Vaccinated with updated booster', 'Vaccinated without updated booster'))+
  scale_x_date(date_breaks= "1 month", date_labels = "%b %Y", limit=as.Date(c('2022-08-01', '2023-03-26')), expand=c(0,0))

```
##Creating an animated plot
```{r}
#Need to delete any NAs for an animated plot to work
tidyUSACOVID <- na.omit(tidyUSACOVID)

ggplot(data=tidyUSACOVID, aes(x=Month, y=value, group=name, color=name))+
  geom_line(size = 1.25)+
  ggtitle("Rates of COVID-19 Deaths by Vaccination Status in Ages 18 and Older")+
  theme_clean()+
  theme(axis.text.x=element_text(angle=60, hjust=1, face = "bold"))+
  theme(axis.title.y=element_text(size=14,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5, lineheight = 0.9))+
  theme(plot.title = element_text(face = "bold", size = 13))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(face = "bold", size = 8.5))+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits=c(0,5))+
  labs(y = "Deaths per 100,000 population", subtitle = "August 1, 2022 - March 26, 2023 (23 U.S. Jurisdictions)")+
  theme(plot.subtitle = element_text(hjust=0.5))+
  scale_color_hue(labels=c('Unvaccinated', 'Vaccinated with updated booster', 'Vaccinated without updated booster'))+
  scale_x_date(date_breaks= "1 month", date_labels = "%b %Y", limit=as.Date(c('2022-08-01', '2023-03-26')), expand=c(0,0))+
  transition_reveal(Month)

```

##Creating a plot for COVID-19 cases by vaccination status
```{r}
USACOVIDCASES <- read.csv ("COVID_Data_R3.csv")
```

```{r}
tidyUSACOVIDCASES <-USACOVIDCASES %>% 
  pivot_longer(cols = -Month)
```

```{r}
tidyUSACOVIDCASES
```
```{r}
tidyUSACOVIDCASES$Month <- mdy(tidyUSACOVIDCASES$Month)
```

```{r}
tidyUSACOVIDCASES
```
```{r}
ggplot(data=tidyUSACOVIDCASES, aes(x=Month, y=value, group=name, color=name))+
  geom_line(size = 1.25)+
  ggtitle("Rates of COVID-19 Cases by Vaccination Status in Ages 18 and Older")+
  theme_clean()+
  theme(axis.text.x=element_text(angle=60, hjust=1, face = "bold"))+
  theme(axis.title.y=element_text(size=14,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5, lineheight = 0.9))+
  theme(plot.title = element_text(face = "bold", size = 14))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(face = "bold", size = 10))+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits=c(0,600))+
  labs(y = "Cases per 100,000 population", subtitle = "August 1, 2022 - April 16, 2023 (24 U.S. Jurisdictions)")+
  theme(plot.subtitle = element_text(hjust=0.5))+
  scale_color_hue(labels=c('Unvaccinated', 'Vaccinated with updated booster', 'Vaccinated without updated booster'))+
  scale_x_date(date_breaks= "1 month", date_labels = "%b %Y", limit=as.Date(c('2022-08-01', '2023-04-16')), expand=c(0,0))
```
##Creating animated plot
```{r}
#Need to delete any NAs for an animated plot to work
tidyUSACOVIDCASES <- na.omit(tidyUSACOVIDCASES)

ggplot(data=tidyUSACOVIDCASES, aes(x=Month, y=value, group=name, color=name))+
  geom_line(size = 1.25)+
  ggtitle("Rates of COVID-19 Cases by Vaccination Status in Ages 18 and Older")+
  theme_clean()+
  theme(axis.text.x=element_text(angle=60, hjust=1, face = "bold"))+
  theme(axis.title.y=element_text(size=14,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5, lineheight = 0.9))+
  theme(plot.title = element_text(face = "bold", size = 13))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(face = "bold", size = 8.5))+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits=c(0,600))+
  labs(y = "Cases per 100,000 population", subtitle = "August 1, 2022 - April 16, 2023 (24 U.S. Jurisdictions)")+
  theme(plot.subtitle = element_text(hjust=0.5))+
  scale_color_hue(labels=c('Unvaccinated', 'Vaccinated with updated booster', 'Vaccinated without updated booster'))+
  scale_x_date(date_breaks= "1 month", date_labels = "%b %Y", limit=as.Date(c('2022-08-01', '2023-04-16')), expand=c(0,0))+
  transition_reveal(Month)

```
##Installing the gifski package in order to save animated plots as gifs
```{r}
install.packages("gifski")
```

```{r}
anim_save("COVID Deaths by Vax Status.gif", animation)
```

##Saving the cases by vax status animated plot for a gif
```{r}
animated_plot <- ggplot(data=tidyUSACOVID, aes(x=Month, y=value, group=name, color=name))+
  geom_line(size = 1.25)+
  ggtitle("Rates of COVID-19 Deaths by Vaccination Status in Ages 18 and Older")+
  theme_clean()+
  theme(axis.text.x=element_text(angle=60, hjust=1, face = "bold"))+
  theme(axis.title.y=element_text(size=14,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5, lineheight = 0.9))+
  theme(plot.title = element_text(face = "bold", size = 13))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(face = "bold", size = 8.5))+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits=c(0,5))+
  labs(y = "Deaths per 100,000 population", subtitle = "August 1, 2022 - March 26, 2023 (23 U.S. Jurisdictions)")+
  theme(plot.subtitle = element_text(hjust=0.5))+
  scale_color_hue(labels=c('Unvaccinated', 'Vaccinated with updated booster', 'Vaccinated without updated booster'))+
  scale_x_date(date_breaks= "1 month", date_labels = "%b %Y", limit=as.Date(c('2022-08-01', '2023-03-26')), expand=c(0,0))+
  transition_reveal(Month)
```

```{r}
COVID_Deaths <- animate(animated_plot)
```

```{r}
anim_save("COVID Deaths by Vax Status.gif", COVID_Deaths)
```

#Saving the cases by vax status animated plot for a gif
```{r}
animated_plot2 <- ggplot(data=tidyUSACOVIDCASES, aes(x=Month, y=value, group=name, color=name))+
  geom_line(size = 1.25)+
  ggtitle("Rates of COVID-19 Cases by Vaccination Status in Ages 18 and Older")+
  theme_clean()+
  theme(axis.text.x=element_text(angle=60, hjust=1, face = "bold"))+
  theme(axis.title.y=element_text(size=14,face="bold"))+
  theme(plot.title = element_text(hjust = 0.5, lineheight = 0.9))+
  theme(plot.title = element_text(face = "bold", size = 13))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(face = "bold", size = 8.5))+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits=c(0,600))+
  labs(y = "Cases per 100,000 population", subtitle = "August 1, 2022 - April 16, 2023 (24 U.S. Jurisdictions)")+
  theme(plot.subtitle = element_text(hjust=0.5))+
  scale_color_hue(labels=c('Unvaccinated', 'Vaccinated with updated booster', 'Vaccinated without updated booster'))+
  scale_x_date(date_breaks= "1 month", date_labels = "%b %Y", limit=as.Date(c('2022-08-01', '2023-04-16')), expand=c(0,0))+
  transition_reveal(Month)
```

```{r}
COVID_Cases <- animate(animated_plot2)
```

```{r}
anim_save("COVID Cases by Vax Status.gif", COVID_Cases)
```

